### 2020-04-04 05-translation.tex
more typos and notation
widetilde widehat proof in appendice

  \draw (4.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
  \draw (5.2,0.255)  node[minimum size=0.485cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  \hat{s1}_i}$};
  \draw(5,1) -- (5,0.5) -- (5.5,0.5) -- (5.5,0);

@@

    \draw (6,1) -- (10,1) ;
 %   \draw (7,1) -- (7,0.5) -- (7.5,0.5) -- (7.5,0);
    \draw (6.5,0) -- (10.5,0);
    \draw   (5.99,1)-- (5.99,0.5) -- (6.5,0.5) -- (6.5,0.5) -- (6.5,0) ;
    \
     \draw (6.75,0.75)  node[minimum size=0.47cm,fill=yellow, draw=yellow] {$\ $};
 \draw (6.36,0.75) node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] 
 {${\scriptscriptstyle \hat{s2}_o \ }$};
 %node[minimum size=0.5cm,draw,draw=white] {$\noact$};
   \draw (6.75,0.26)  node[minimum size=0.485cm,fill=yellow, draw=yellow] {$\ $};

>>>>

  \draw (4.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
  \draw (5.2,0.255)  node[minimum size=0.485cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  \widehat{s1}_i}$};
  \draw(5,1) -- (5,0.5) -- (5.5,0.5) -- (5.5,0);

@@

    \draw (6,1) -- (10,1) ;
 %   \draw (7,1) -- (7,0.5) -- (7.5,0.5) -- (7.5,0);
    \draw (6.5,0) -- (10.5,0);
    \draw   (5.99,1)-- (5.99,0.5) -- (6.5,0.5) -- (6.5,0.5) -- (6.5,0) ;
    \
     \draw (6.75,0.75)  node[minimum size=0.47cm,fill=yellow, draw=yellow] {$\ $};
 \draw (6.36,0.75) node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] 
 {${\scriptscriptstyle \widehat{s2}_o \ }$};
 %node[minimum size=0.5cm,draw,draw=white] {$\noact$};
   \draw (6.75,0.26)  node[minimum size=0.485cm,fill=yellow, draw=yellow] {$\ $};

### 2020-04-04 06b-bisimulation.tex
widetilde widehat proof in appendice

\begin{enumerate}
\item
$\star :: r_j :: [ s_1,...,s_n ]^* :: s_i :: [  s_1,...,s_n ]^* :: [  \overline{s}_1,...,\overline{s}_n ]^+ :: \star$
\item
$\star :: r_j :: [s_i, \overline{s}_i] :: r_{j+1} :: \star$
\item
$\star :: p_j :: [  \mathit{ents} ]^* :: \tilde{s}_i :: \star$
\item
$\star :: \tilde{s}_i :: \star$
\item
$\star :: \hat{s}_i :: \star$
\item
$\star :: r_j :: ? :: r_{j+1} :: \star$
\end{enumerate}
where in 1, 2, 6 we exploit the fact that in a reaction $(R,I,P)$ the sets $R$ of reactants and  $I$ of inhibitors are non empty, so that if there is only one symbol between the occurrence of $r_j$ and $r_{j+1}$ it means the reaction $r_j$ has not been applied. Viceversa, if the reaction $r_j$ has been applied the occurrence of $r_j$ must be followed by at least one of the symbols in $\{s_1,...,s_n\}$ and then by at least one of the symbols in $\{\overline{s}_1,...,\overline{s}_n\}$.

@@

\begin{theorem}
$\sim_F\ =\ \equiv_{\mathcal{L}_F}$
\end{theorem}
\begin{proof}
The proof is just an adaptation of the classical result. The two implications are proved separately.
\begin{description}
\item[$\sim_F\ \subseteq\ \equiv_{\mathcal{L}_F}$:]
Given any two processes $P\sim_F Q$ we need to prove that for any bioHML formula $G$ we have $P\entails G$ iff $Q\entails G$. 
Without loss of generality, we prove that $P\entails G$ implies $Q\entails G$.
The proof is by structural induction on $G$.
\begin{itemize}
\item 
if $G={\tt t}$, then $Q\entails G$.

\item
if $G={\tt f}$, then the assumption $P\entails G$ is false and the implication holds.

\item
if $G=G_1 \wedge G_2$ we take as inductive hypotheses that
$$
\begin{array}{l}
\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
\end{array}
$$
We need to prove that $Q\entails G$.
Since $P\entails G=G_1 \wedge G_2$ we have $P\entails G_1$ and $P\entails G_2$.
Since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and $Q\entails G_2$.
Hence $Q\entails G_1 \wedge G_2=G$.

\item
if $G=G_1 \vee G_2$  we take as inductive hypotheses that
$$
\begin{array}{l}
\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
\end{array}
$$
We need to prove that $Q\entails G$.
Since $P\entails G=G_1 \vee G_2$ we have $P\entails G_1$ or $P\entails G_2$.
If $P\entails G_1$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and thus $Q\entails G_1 \vee G_2=G$.
If $P\entails G_2$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_2$ and thus $Q\entails G_1 \vee G_2=G$.

\item
if $G=\langle\chi\rangle H$ we take as inductive hypothesis that
$$
\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
$$
We need to prove that $Q\entails G$.
Since $P\entails \langle\chi\rangle H$ it means that there exists $\upsilon, P'$ such that $P\xRightarrow{\upsilon} P'$ with $\upsilon\entails \chi$ and $P'\entails H$.
Since $P\sim_F Q$, there exists $w,Q'$ such that $Q\xRightarrow{w} Q'$ with $w\entails \chi$ and $P'\sim_F Q'$. Then, by inductive hypothesis, $Q'\entails H$ and thus $Q\entails \langle\chi\rangle H=G$.

\item
if $G=[\chi] H$ we take as inductive hypothesis that
$$
\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
$$
We need to prove that $Q\entails G$.
If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$ then 
 $Q\entails [\chi] H=G$ trivially.
For any $\upsilon,Q'$ such that $Q\xRightarrow{\upsilon}Q'$ with $\upsilon\entails \chi$, then as $P\sim_F Q$ there must exist $w,P'$ such that $P\xRightarrow{w}P'$ with $w\entails \chi$ and $P'\sim_F Q'$. Since $P\entails G=[\chi] H$ then it must be $P'\entails H$. Since $P'\sim_F Q'$, by inductive hypothesis $Q'\entails H$. Hence $Q\entails [\chi] H = G$.
\end{itemize}

\item[$\equiv_{\mathcal{L}_F}\ \subseteq\ \sim_F$:]
We prove that $\equiv_{\mathcal{L}_F}$ is a bio-simulation and thus included in $\sim_F$.
Take two generic processes $P\equiv_{\mathcal{L}_F} Q$ and suppose $P\xRightarrow{\upsilon}P'$ for some $\upsilon,P'$. 
\begin{itemize}
\item
If $\upsilon \entails F$ we want to prove that there exists some $w,Q'$ such that $Q\xRightarrow{w}Q'$, with $w\entails F$ and $P'\equiv_{\mathcal{L}_F} Q'$.

Towards a contradiction, assume that we cannot find such $w,Q'$.

If there is no transition $Q\xRightarrow{w}Q'$ such that $w\entails F$, then the bioHML formula $G \defeq \langle F\rangle{\tt t}$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.

Otherwise, let $\mathcal{Q}\defeq \{Q'~\mid~\exists w.\ Q\xRightarrow{w}Q' \ \wedge\ w\entails F\}$ be the set of processes reachable from $Q$ via a transition with a (complete) label that satisfies $F$. Since our processes are with guarded recursion, the set $\mathcal{Q}$ is finite. Let $\mathcal{Q} = \{Q'_1,...,Q'_n\}$. By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $G_i$ such that $P'\entails G_i$ and $Q'_i\not\entails G_i$ (if it was the opposite, $P'\not\entails G_i$ and $Q'_i\entails G_i$, we can use the converse formulas). But then the formula $G\defeq \langle F\rangle(G_1\wedge\dots\wedge G_n)$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.

\item
If $\upsilon \not\entails F$ then the proof is analogous to the previous case and thus omitted.
\end{itemize}
\end{description}
\end{proof}

>>>>

\begin{enumerate}
\item
$\star :: r_j :: [ s_1,...,s_n ]^* :: s_i :: [  s_1,...,s_n ]^* :: [  \overline{s}_1,...,\overline{s}_n ]^+ :: \star$
\item
$\star :: r_j :: [s_i, \overline{s}_i] :: r_{j+1} :: \star$
\item
$\star :: p_j :: [  \mathit{ents} ]^* :: \widetilde{s}_i :: \star$
\item
$\star :: \widetilde{s}_i :: \star$
\item
$\star :: \widehat{s}_i :: \star$
\item
$\star :: r_j :: ? :: r_{j+1} :: \star$
\end{enumerate}
where in 1, 2, 6 we exploit the fact that in a reaction $(R,I,P)$ the sets $R$ of reactants and  $I$ of inhibitors are non empty, so that if there is only one symbol between the occurrence of $r_j$ and $r_{j+1}$ it means the reaction $r_j$ has not been applied. Viceversa, if the reaction $r_j$ has been applied the occurrence of $r_j$ must be followed by at least one of the symbols in $\{s_1,...,s_n\}$ and then by at least one of the symbols in $\{\overline{s}_1,...,\overline{s}_n\}$.

@@

\begin{restatable}[Correspondence]{theorem}{theocorr}
%\begin{theorem}
$\sim_F\ =\ \equiv_{\mathcal{L}_F}$
%\end{theorem}
\end{restatable}
%\begin{proof}
%The proof is just an adaptation of the classical result. The two implications are proved separately.
%\begin{description}
%\item[$\sim_F\ \subseteq\ \equiv_{\mathcal{L}_F}$:]
%Given any two processes $P\sim_F Q$ we need to prove that for any bioHML formula $G$ we have $P\entails G$ iff $Q\entails G$. 
%Without loss of generality, we prove that $P\entails G$ implies $Q\entails G$.
%The proof is by structural induction on $G$.
%\begin{itemize}
%\item 
%if $G={\tt t}$, then $Q\entails G$.
%
%\item
%if $G={\tt f}$, then the assumption $P\entails G$ is false and the implication holds.
%
%\item
%if $G=G_1 \wedge G_2$ we take as inductive hypotheses that
%$$
%\begin{array}{l}
%\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
%\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
%\end{array}
%$$
%We need to prove that $Q\entails G$.
%Since $P\entails G=G_1 \wedge G_2$ we have $P\entails G_1$ and $P\entails G_2$.
%Since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and $Q\entails G_2$.
%Hence $Q\entails G_1 \wedge G_2=G$.
%
%
%\item
%if $G=G_1 \vee G_2$  we take as inductive hypotheses that
%$$
%\begin{array}{l}
%\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
%\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
%\end{array}
%$$
%We need to prove that $Q\entails G$.
%Since $P\entails G=G_1 \vee G_2$ we have $P\entails G_1$ or $P\entails G_2$.
%If $P\entails G_1$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and thus $Q\entails G_1 \vee G_2=G$.
%If $P\entails G_2$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_2$ and thus $Q\entails G_1 \vee G_2=G$.
%
%\item
%if $G=\langle\chi\rangle H$ we take as inductive hypothesis that
%$$
%\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
%$$
%We need to prove that $Q\entails G$.
%Since $P\entails \langle\chi\rangle H$ it means that there exists $\upsilon, P'$ such that $P\xRightarrow{\upsilon} P'$ with $\upsilon\entails \chi$ and $P'\entails H$.
%Since $P\sim_F Q$, there exists $w,Q'$ such that $Q\xRightarrow{w} Q'$ with $w\entails \chi$ and $P'\sim_F Q'$. Then, by inductive hypothesis, $Q'\entails H$ and thus $Q\entails \langle\chi\rangle H=G$.
%
%\item
%if $G=[\chi] H$ we take as inductive hypothesis that
%$$
%\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
%$$
%We need to prove that $Q\entails G$.
%If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$, then 
% $Q\entails [\chi] H=G$ trivially.
%For any $\upsilon,Q'$ such that $Q\xRightarrow{\upsilon}Q'$ with $\upsilon\entails \chi$, then as $P\sim_F Q$ there must exist $w,P'$ such that $P\xRightarrow{w}P'$ with $w\entails \chi$ and $P'\sim_F Q'$. Since $P\entails G=[\chi] H$ then it must be $P'\entails H$. Since $P'\sim_F Q'$, by inductive hypothesis $Q'\entails H$. Hence $Q\entails [\chi] H = G$.
%\end{itemize}
%
%\item[$\equiv_{\mathcal{L}_F}\ \subseteq\ \sim_F$:]
%We prove that $\equiv_{\mathcal{L}_F}$ is a bio-simulation and thus included in $\sim_F$.
%Take two generic processes $P\equiv_{\mathcal{L}_F} Q$ and suppose $P\xRightarrow{\upsilon}P'$ for some $\upsilon,P'$. 
%\begin{itemize}
%\item
%If $\upsilon \entails F$ we want to prove that there exists some $w,Q'$ such that $Q\xRightarrow{w}Q'$, with $w\entails F$ and $P'\equiv_{\mathcal{L}_F} Q'$.
%
%Towards a contradiction, assume that we cannot find such $w,Q'$.
%
%If there is no transition $Q\xRightarrow{w}Q'$ such that $w\entails F$, then the bioHML formula $G \defeq \langle F\rangle{\tt t}$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
%
%Otherwise, let $\mathcal{Q}\defeq \{Q'~\mid~\exists w.\ Q\xRightarrow{w}Q' \ \wedge\ w\entails F\}$ be the set of processes reachable from $Q$ via a transition with a (complete) label that satisfies $F$. Since our processes are with guarded recursion, the set $\mathcal{Q}$ is finite. Let $\mathcal{Q} = \{Q'_1,...,Q'_n\}$. By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $G_i$ such that $P'\entails G_i$ and $Q'_i\not\entails G_i$ (if it was the opposite, $P'\not\entails G_i$ and $Q'_i\entails G_i$, we can use the converse formulas). But then the formula $G\defeq \langle F\rangle(G_1\wedge\dots\wedge G_n)$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
%
%\item
%If $\upsilon \not\entails F$ then the proof is analogous to the previous case and thus omitted.
%\end{itemize}
%\end{description}
%\end{proof}

### 2020-04-04 07-discussion.tex
widetilde widehat proof in appendice

Our encoding increases the expressivity of  RS  concerning:
the behaviour of the context, the possibility of alternative behaviour of
mutated entities, the communication between two different reaction systems.
It is important to note that our encoding guarantees that from each state, 
in the  c\CNA \ transition system, only one transition comes out, 
as the dynamics is totally deterministic.

\subsection{Recursive contexts}
In RS, the behaviour of the context is finite. For the first $n$ steps, it is specified 
 which are the entities that are provided from the context. 
Using c\CNA\ we can describe in a natural way the behaviour of the context in a recursive way. 
Then, the context behaviour would not necessarily end after $n$ steps,
and  could be infinite.
For example, in an extended interactive process, we may want that the entity $s$ is 
intermittently provided by the context every two steps: 
\[
\begin{array}{lcll}
Cxt_s & \defeq &  \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}.Cxt_{s'}, &\mbox{ context provides $s$;} \\
Cxt_{s'} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s''},& \mbox{ context doesn't provide $s$;} \\
Cxt_{s''} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s}, & \mbox{ context doesn't provide $s$.} \\
\end{array}
\]

>>>>

Our encoding increases the expressivity of  RS  concerning:
%the behaviour of the context, 
the possibility of alternative behaviour of
mutated entities, and  the communication between two different reaction systems.
It is important to note that our encoding guarantees that from each state, 
in the  c\CNA \ transition system, only one transition comes out, 
as the dynamics is totally deterministic.

%\subsection{Recursive contexts}
%In RS, the behaviour of the context is finite. For the first $n$ steps, it is specified 
% which are the entities that are provided from the context. 
%Using c\CNA\ we can describe in a natural way the behaviour of the context in a recursive way. 
%Then, the context behaviour would not necessarily end after $n$ steps,
%and  could be infinite.
%For example, in an extended interactive process, we may want that the entity $s$ is 
%intermittently provided by the context every two steps: 
%\[
%\begin{array}{lcll}
%Cxt_s & \defeq &  \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}.Cxt_{s'}, &\mbox{ context provides $s$;} \\
%Cxt_{s'} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s''},& \mbox{ context doesn't provide $s$;} \\
%Cxt_{s''} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s}, & \mbox{ context doesn't provide $s$.} \\
%\end{array}
%\]

### 2020-04-04 08-conclusion.tex
widetilde widehat proof in appendice

In this paper we have introduced a variant of the {\tt link}-calculus
where prefixes are link chains and no more single links, as it was 
briefly described in the future work section in~\cite{BBB17}.
This variant allowed us to define 
%shown 
an elegant embedding of reaction systems,
an emerging formalism to model computationally biochemical systems.
%into multilink processes.  
This translation shows several benefits.
For instance, the context behaviour can also be expressed recursively. 
Entity mutations can be expressed easily. Reaction systems can
communicate between them.

>>>>

In this paper we have introduced a variant of the {\tt link}-calculus
where prefixes are link chains and no more single links, as it was 
briefly described in the future work section in~\cite{BBB17}.
This variant allowed us to define 
%shown 
an elegant embedding of reaction systems,
an emerging formalism to model computationally biochemical systems.
%into multilink processes.  
This translation shows several benefits.
{\color{red} For instance, in our paper the context works non deterministically and it is recursively defined.}
%  the context behaviour is expressed recursively and no be expressed recursively. 
Also, entity mutations can be expressed easily and different reaction systems can
communicate between them.

### 2020-04-04 09-appendix.tex
widetilde widehat proof in appendice

\section{Omitted Proofs}
In this section we report the proofs for the results in Section~\ref{sec:trans}.
%\noindent
%We recall Lemma~\ref{lem:struct} (the original appears on p.\,\pageref{lem:struct}).

@@

% \begin{corollary}[completeness2]
% \label{prop:completeness}
%  Let ${\cal A} = (S, A)$ be a $rs$, and let $\pi=(\gamma,\delta)$ be a $n$-step interactive process in ${\cal A}$ with $\gamma = C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$. Let $P=[|{\cal A},\gamma|]$ its c\CNA \ translation. If we define $\pi'= (\gamma',\delta')$  with   $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$  as
%  $C_0'=C_w \cup D_w$, and   $C_i'=C_{i+w}$, $D_i' = D_{i+w}$ with $1 \leq i \leq n-w $, then $\exists$ $P'$ such that $P \rightarrow^w P'$ and  $P'=[|{\cal A},\gamma'|]$.  
% \end{corollary}
 \begin{proof}
 The proof proceeds by induction on the number $j$, and it is similar to the one of Corollary~\ref{corr:corr2}\\
% case $j=0$\\
% It is enough to set $C_0' = C_0 \cup D_0$ and $P' =P$.\\
% case $j=1$\\
% This case falls into the case of Proposition~\ref{prop:compl1}.\\
% case $j>1$\\
% We set $C_0'=C_w \cup D_w = C_1'' \cup D_1''$, $C_g' = C_{n-w}= C_{g+1}''$, and $D_g' = D_{n-w} = D_{g+1}''$, with $1 \leq  g < n-w$.\\
% By inductive hypothesis, it holds that $\exists$ $P''$ such that $P \xrightarrow{\link{\silent}{\silent}}^{w-1} P''$ and $P'' = [| {\cal A},\gamma'|]$.
% By Lemma~\ref{lem:compl2}.1, it exists $P'$ such that $P'' \xrightarrow{\link{\silent}{\silent}} P'$,
% and $P' = [|{\cal A},\gamma'|]$. By Lemma~\ref{lem:compl2}.2, $P'$ is unique.
%% By Lemma~\ref{lem:compl2}.2, if $\exists$ $P'$ such that $P'' \xrightarrow{\link{\silent}{\silent}} P'$, then $P'$ is 
% %unique, and by Lemma~\ref{lem:compl2}.1 $P' = [|{\cal A},\gamma'|]$. 
% And we are done. 
 \end{proof}

>>>>

\section{Omitted Proofs}
In this section we report the proofs for the results in Section~\ref{sec:trans} and in Section~\ref{sec:biosimulation}.
%\noindent
%We recall Lemma~\ref{lem:struct} (the original appears on p.\,\pageref{lem:struct}).

@@

% \begin{corollary}[completeness2]
% \label{prop:completeness}
%  Let ${\cal A} = (S, A)$ be a $rs$, and let $\pi=(\gamma,\delta)$ be a $n$-step interactive process in ${\cal A}$ with $\gamma = C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$. Let $P=[|{\cal A},\gamma|]$ its c\CNA \ translation. If we define $\pi'= (\gamma',\delta')$  with   $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$  as
%  $C_0'=C_w \cup D_w$, and   $C_i'=C_{i+w}$, $D_i' = D_{i+w}$ with $1 \leq i \leq n-w $, then $\exists$ $P'$ such that $P \rightarrow^w P'$ and  $P'=[|{\cal A},\gamma'|]$.  
% \end{corollary}
 \begin{proof}
 The proof proceeds by induction on the number $j$, and it is similar to the one of Corollary~\ref{corr:corr2}.\\
% case $j=0$\\
% It is enough to set $C_0' = C_0 \cup D_0$ and $P' =P$.\\
% case $j=1$\\
% This case falls into the case of Proposition~\ref{prop:compl1}.\\
% case $j>1$\\
% We set $C_0'=C_w \cup D_w = C_1'' \cup D_1''$, $C_g' = C_{n-w}= C_{g+1}''$, and $D_g' = D_{n-w} = D_{g+1}''$, with $1 \leq  g < n-w$.\\
% By inductive hypothesis, it holds that $\exists$ $P''$ such that $P \xrightarrow{\link{\silent}{\silent}}^{w-1} P''$ and $P'' = [| {\cal A},\gamma'|]$.
% By Lemma~\ref{lem:compl2}.1, it exists $P'$ such that $P'' \xrightarrow{\link{\silent}{\silent}} P'$,
% and $P' = [|{\cal A},\gamma'|]$. By Lemma~\ref{lem:compl2}.2, $P'$ is unique.
%% By Lemma~\ref{lem:compl2}.2, if $\exists$ $P'$ such that $P'' \xrightarrow{\link{\silent}{\silent}} P'$, then $P'$ is 
% %unique, and by Lemma~\ref{lem:compl2}.1 $P' = [|{\cal A},\gamma'|]$. 
% And we are done. 
 \end{proof}

 \theocorr*
 \begin{proof}
The proof is just an adaptation of the classical result. The two implications are proved separately.
\begin{description}
\item[$\sim_F\ \subseteq\ \equiv_{\mathcal{L}_F}$:]
Given any two processes $P\sim_F Q$ we need to prove that for any bioHML formula $G$ we have $P\entails G$ iff $Q\entails G$. 
Without loss of generality, we prove that $P\entails G$ implies $Q\entails G$.
The proof is by structural induction on $G$.
\begin{itemize}
\item 
if $G={\tt t}$, then $Q\entails G$.

\item
if $G={\tt f}$, then the assumption $P\entails G$ is false and the implication holds.

\item
if $G=G_1 \wedge G_2$ we take as inductive hypotheses that
$$
\begin{array}{l}
\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
\end{array}
$$
We need to prove that $Q\entails G$.
Since $P\entails G=G_1 \wedge G_2$ we have $P\entails G_1$ and $P\entails G_2$.
Since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and $Q\entails G_2$.
Hence $Q\entails G_1 \wedge G_2=G$.

\item
if $G=G_1 \vee G_2$  we take as inductive hypotheses that
$$
\begin{array}{l}
\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
\end{array}
$$
We need to prove that $Q\entails G$.
Since $P\entails G=G_1 \vee G_2$ we have $P\entails G_1$ or $P\entails G_2$.
If $P\entails G_1$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and thus $Q\entails G_1 \vee G_2=G$.
If $P\entails G_2$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_2$ and thus $Q\entails G_1 \vee G_2=G$.

\item
if $G=\langle\chi\rangle H$ we take as inductive hypothesis that
$$
\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
$$
We need to prove that $Q\entails G$.
Since $P\entails \langle\chi\rangle H$ it means that there exists $\upsilon, P'$ such that $P\xRightarrow{\upsilon} P'$ with $\upsilon\entails \chi$ and $P'\entails H$.
Since $P\sim_F Q$, there exists $w,Q'$ such that $Q\xRightarrow{w} Q'$ with $w\entails \chi$ and $P'\sim_F Q'$. Then, by inductive hypothesis, $Q'\entails H$ and thus $Q\entails \langle\chi\rangle H=G$.

\item
if $G=[\chi] H$ we take as inductive hypothesis that
$$
\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
$$
We need to prove that $Q\entails G$.
If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$, then 
 $Q\entails [\chi] H=G$ trivially.
For any $\upsilon,Q'$ such that $Q\xRightarrow{\upsilon}Q'$ with $\upsilon\entails \chi$, then as $P\sim_F Q$ there must exist $w,P'$ such that $P\xRightarrow{w}P'$ with $w\entails \chi$ and $P'\sim_F Q'$. Since $P\entails G=[\chi] H$ then it must be $P'\entails H$. Since $P'\sim_F Q'$, by inductive hypothesis $Q'\entails H$. Hence $Q\entails [\chi] H = G$.
\end{itemize}

\item[$\equiv_{\mathcal{L}_F}\ \subseteq\ \sim_F$:]
We prove that $\equiv_{\mathcal{L}_F}$ is a bio-simulation and thus included in $\sim_F$.
Take two generic processes $P\equiv_{\mathcal{L}_F} Q$ and suppose $P\xRightarrow{\upsilon}P'$ for some $\upsilon,P'$. 
\begin{itemize}
\item
If $\upsilon \entails F$ we want to prove that there exists some $w,Q'$ such that $Q\xRightarrow{w}Q'$, with $w\entails F$ and $P'\equiv_{\mathcal{L}_F} Q'$.

Towards a contradiction, assume that we cannot find such $w,Q'$.

If there is no transition $Q\xRightarrow{w}Q'$ such that $w\entails F$, then the bioHML formula $G \defeq \langle F\rangle{\tt t}$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.

Otherwise, let $\mathcal{Q}\defeq \{Q'~\mid~\exists w.\ Q\xRightarrow{w}Q' \ \wedge\ w\entails F\}$ be the set of processes reachable from $Q$ via a transition with a (complete) label that satisfies $F$. Since our processes are with guarded recursion, the set $\mathcal{Q}$ is finite. Let $\mathcal{Q} = \{Q'_1,...,Q'_n\}$. By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $G_i$ such that $P'\entails G_i$ and $Q'_i\not\entails G_i$ (if it was the opposite, $P'\not\entails G_i$ and $Q'_i\entails G_i$, we can use the converse formulas). But then the formula $G\defeq \langle F\rangle(G_1\wedge\dots\wedge G_n)$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.

\item
If $\upsilon \not\entails F$ then the proof is analogous to the previous case and thus omitted.
\end{itemize}
\end{description}
\end{proof}
